\documentclass{fast-nuces-bs}

\usepackage{blindtext}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage{lipsum} 
\usepackage{sectsty}
\usepackage{tikz-uml}
\usepackage{titlesec}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{float}

% Code listing settings
\lstset{
    basicstyle=\ttfamily\small,
    breaklines=true,
    frame=single,
    numbers=left,
    numberstyle=\tiny,
    backgroundcolor=\color{gray!10}
}

% Information about the Project
\department{Department of Computer Science}
\faculty{Computer Science}
\degreeyear{2026}
\degreemonth{May}
\degreename{Computer Science}
\campuscity{Islamabad}
\authorone{Ali Irfan}{i212572}
\authortwo{Rafay Khattak}{i210423}
\authorthree{Nouman Hafeez}{i210416}
\supervisor{Ms. Kainat Iqbal}
\sessionduration{2021-2026}
\cosupervisor{Ms. Saira Qamar}
\deanname{Dr. Dean Computing}
\directorname{Dr. Director of the Campus}
\hodname{Dr. Head of Computer Science Department}
\fypcoordinatorname{FYP Coordinator}
\title{CORAL: Consensus-based Refinement And Learning - A Multi-Hypothesis Correction Architecture for State-of-the-Art Urdu ASR - Iteration 1 Report}

\begin{document}

\begin{abstract}
This report presents the outcomes of Iteration 1 for the CORAL (Consensus-based Refinement And Learning) project, a novel multi-hypothesis ASR system for low-resource Urdu language. The primary objective of this iteration was to establish the foundational infrastructure by integrating an ensemble of pre-trained ASR models and implementing word-level confidence score extraction mechanisms. We successfully deployed four state-of-the-art models (Whisper Large, Whisper Medium, Whisper Small, and Wav2Vec2-XLSR-Urdu) and evaluated their baseline performance on a 10-sample test set from the Common Voice Urdu dataset. Results demonstrate that Whisper Large achieves the best baseline WER of 17.76\%, establishing a performance benchmark for subsequent iterations. The complete pipeline for confidence-annotated hypothesis generation is now operational, with comprehensive evaluation metrics including WER, CER, confidence scores, and Expected Calibration Error (ECE). This foundational work enables the development of the instruction-guided correction system in Iteration 2.
\end{abstract}

\include{sections/chapter1}
\include{sections/chapter2}
\include{sections/chapter3}

\bibliographystyle{plain}
\bibliography{bib} 
\addcontentsline{toc}{chapter}{References} 

\end{document}